---
title: "Assignment 11"
author: "James Dethlefs"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Import Data
```{r}
raw0 <- read.csv(file = "State_Employee_Payroll_Data_Calendar_Year_2015_through_Present.tsv", sep = "\t", header = TRUE)

```

###Data Wrangling
```{r, echo=FALSE}
library(dplyr)
```


```{r}
employID <- unique(raw0$EmplId.Empl.Rcd)

# pick full time employee
raw1 <- filter(raw0, Full.Part == "F")  
# keep only the interested variables
raw2 <- raw1[, -c(1, 5:9, 11:12, 19:23, 31:33, 35:36)] 

EmplId.Empl.Rcd <- as.character(raw2$EmplId.Empl.Rcd)
counts <- data.frame(table(EmplId.Empl.Rcd))
PersonOccurCount <- table(counts[, 2])
plot(PersonOccurCount, col = rainbow(30), 
     xlab = "Occurance of employee", ylab = "Count of employee")
```

```{r}
#Remove time stamp of date
raw2$Check.Dt <- gsub(" 12:00:00 AM", "", raw2$Check.Dt)

#Remove 'U' Sex
raw.sex <- filter(raw2, raw2$Sex != 'U')
```



#Figure 1
Distribution of sex 
```{r}
library(ggplot2)
library(scales)

ggplot(raw.sex, aes(x=raw.sex$Sex, fill = raw.sex$Ethnic.Grp)) +
  labs(x = "Gender", y ="Frequency of Gender", fill = "Ethnic Group") +
  geom_bar() + 
  scale_y_continuous(labels = number_format(decimal.mark = ",")) 
      

```

We can see that there are more men in the data than women. There are also more people that identify as white than any other ethnicity. Black's are then next highest ethnicity. The largest subgroup looks to be white females.

#Figure 2
Distribution of Income
```{r}

annual.sum <- sum(raw.sex$Annual.Rate)


ggplot(raw.sex, aes(x = raw.sex$Check.Dt, y = raw.sex$Annnual.Rate, fill = raw.sex$Sex)) + 
  labs(x = "Date", y = "Annual Rate", fill = "Gender") +
  geom_line() +
  facet_grid(raw.sex$Calendar.Year) 

```

Although the plots are hard to read and get specific's from, but we can see that the annual rate has decreasedas the years have gone on. We also see some seasonality in each year


#Figure 3
```{r}

race.sum <- aggregate(data = raw.sex, raw.sex$Annnual.Rate ~ raw.sex$Ethnic.Grp, sum)

barplot(race.sum[,2], xlab = "Ethnic Group", ylab = "Sum of each Ethinic Group", names.arg= race.sum[,1], las = 2)
```

